---
title: "Homework 5"
author: "Mengfan Luo"
date: "11/20/2021"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 1,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1

### (a) Data loading, description and summary

```{r}
homicide_data = read_csv("data/homicide-data.csv",na = c("","Unknown"))
```

In the raw data, there are ``r nrow(homicide_data)`` observations and ``r ncol(homicide_data)`` columns. There are `r sum(is.na(homicide_data))` missing values, including 60 lying in variable `lat`, 60 in `lon`, and 2999 in `victim_age`. The variables include 8 character variables `uid`,`victim_last`,`victim_first`,`victim_race`,`victim_sex`,`city`,`state` and `disposition	`, and 4 numeric variables `reported_date`,`victim_age`, `reported_date` and `reported_date`.

`city_state` variable is created to display both city and state information. Total number of homicides and the number of unsolved homicides within a city is summarized in the following table. We notice that "Tulsa,AL" only has 1 total case, and in fact Tulsa is a city in OK rather than AL, which indicate this might be a error in data entry. We remove row with `city_state == "Tulsa,AL"` and keep `city_state == "Tulsa,OK"`.

```{r}
homicide_data = homicide_data %>% 
  mutate(city_state = str_c(city,state,sep = ","))

city_summary = homicide_data %>% 
  mutate(
    unsolved = case_when(
      disposition == "Closed by arrest" ~ 0,
      disposition != "Closed by arrest" ~ 1)
    ) %>% 
  group_by(city_state) %>% 
  summarize(city_total = n(),
            unsolved_total = sum(unsolved)) %>% 
  filter(city_state != "Tulsa,AL")
knitr::kable(city_summary)
```

### (b) Proportion of unsolved homicides in Baltimore, MD

For the city of Baltimore, MD,  `prop.test` function is used to estimate the proportion of homicides that are unsolved and the output is saved in `baltimore_output`.

* My initial approach:

```{r}
baltimore_total = city_summary %>%
  filter(city_state == "Baltimore,MD") %>% 
  pull(city_total)

baltimore_unsolved = city_summary %>%
  filter(city_state == "Baltimore,MD") %>% 
  pull(unsolved_total)


baltimore_output = prop.test(baltimore_unsolved,baltimore_total)


baltimore_estimate = broom::tidy(baltimore_output) %>% 
  pull(estimate)

baltimore_conf.low = broom::tidy(baltimore_output) %>% 
  pull(conf.low)

baltimore_conf.high = broom::tidy(baltimore_output) %>% 
  pull(conf.high)

baltimore_CI = c(baltimore_conf.low,baltimore_conf.high)
```

Using broom::tidy on `baltimore_output` to pull the estimated proportion and confidence intervals, we get the estimate is `r baltimore_estimate` and 95% CI is [`r baltimore_CI`].

The proportion of unsolved homicides and the confidence interval for each cities are retrived using the following pipeline, and is shown in the following table.


* Revised method:

```{}
baltimore_summary= city_summary %>% 
  filter(city_state == "Baltimore,MD")

baltimore_test = prop.test(
  x = baltimore_summary %>% pull(unsolved_total),
  n = baltimore_summary %>% pull(city_total)
)

baltimore_test %>% 
  broom::tidy()
```


### (c) Proportion of unsolved homicides in all cities

* My approach

```{r}
city_summary %>% 
  filter(city_state != "Tulsa,AL") %>% 
  mutate(output = map2(unsolved_total,city_total,prop.test),
         output = map(output,broom::tidy)) %>% 
  select(city_state,output) %>% 
  unnest(cols = c(output)) %>% 
  mutate(prop_unsolved = round(estimate,4),
    CI = str_c("[",round(conf.low,4),",",round(conf.high,4),"]")) %>% 
  select(city_state,prop_unsolved,CI) %>% 
  knitr::kable()

```


* Revised approach

```{}
prop_test_function = function(homicide_summary){
  homicide_test = prop.test(
    x = homicide_summary %>% pull(unsolved_total),
    n = homicide_summary %>% pull(city_total)
  )
  return(homicide_test)
}

results_df = city_summary %>%
  nest(data = city_total:unsolved_total) %>% 
  mutate(
    test_results = map(data,prop_test_function),
    tidy_results = map(test_results,broom::tidy)
  ) %>% 
  select(city_state,tidy_results) %>% 
  unnest(tidy_results) %>% 
  select(city_state,estimate,conf.low, conf.high)
  
```

### (d) Visualization of estimatied proportion and CIs of unsolved homicides in all cities

Create a plot that shows the estimates and CIs for each city – check out geom_errorbar for a way to add error bars based on the upper and lower limits. Organize cities according to the proportion of unsolved homicides.

* My approach

```{r}
city_summary %>% 
  mutate(output = map2(unsolved_total,city_total,prop.test),
         output = map(output,broom::tidy)) %>% 
  select(city_state,output) %>% 
  unnest(cols = c(output)) %>% 
  select(city_state,estimate,conf.low,conf.high) %>% 
  mutate(city_state = fct_reorder(city_state,estimate)) %>% 
  ggplot(aes(x = estimate,y = city_state))+
  geom_point()+
  geom_errorbar(aes(xmin = conf.low, xmax = conf.high))+
  labs(
    title = "Estimatied proportion and CIs of unsolved homicides in all cities",
    x = "Estimatied proportion of unsolved homicides",
    y = "City and State"
  )
  
```

* Alternative approach

```{}
city_summary %>% 
  mutate(output = map2(unsolved_total,city_total,prop.test),
         output = map(output,broom::tidy)) %>% 
  select(city_state,output) %>% 
  unnest(cols = c(output)) %>% 
  select(city_state,estimate,conf.low,conf.high) %>% 
  mutate(city_state = fct_reorder(city_state,estimate)) %>% 
  ggplot(aes(x = city_state,y = estimate))+
  geom_point()+
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high))+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```


## Problem 2

### (a) Data loading and tidying

The following dataframe containing data from all participants, including the subject ID, arm, and observations over week1 to week 8.

```{r, message=FALSE}
dataframe = tibble(
  filename = list.files(path = "./data/problem2/")
)

dataframe =  dataframe %>% 
  mutate(data = map(str_c("data/problem2/",filename),read_csv),
         filename = str_remove(filename,".csv")) %>% 
  unnest(data) %>% 
  separate(filename,into = c("arm","ID"),sep = "_") %>% 
  relocate(ID)

knitr::kable(dataframe)
```

### (b) Spaghetti plot of observations on each subject in control arms and experimental arms over weeks

```{r}
dataframe %>% 
  pivot_longer(week_1:week_8,names_to = "week", values_to = "values") %>% 
  mutate(week = str_remove(week,"week_")) %>% 
  ggplot(aes(x = week, y = values, group = ID, color = ID))+
  geom_point(alpha = .5)+
  geom_line(size = .5)+
  facet_grid(.~arm)+
  labs(x = "Week", y = "Values",
       title = "Spaghetti plot of observations on each subject\nin control arms and experimental arms over weeks",
       color = "Subject ID in each group")
```

From the above spaghetti plot, there seem to be a larger increase over week in the experimental arm than in the control arm. The values in week 1 is similar in both groups, but in week 8, all participants in experimental arms have larger values than those in control arms.


## Problem 3

### Data loading, name cleaning and dataset overview

```{r}
set.seed(10)

iris_with_missing = iris %>% 
  map_df(~replace(.x, sample(1:150, 20), NA)) %>%
  mutate(Species = as.character(Species))


iris_with_missing = iris_with_missing %>% 
  janitor::clean_names() 

#skimr::skim(iris_with_missing)

```

Using `skim()`, we can see there are 150 rows ad 5 columns, including 4 numeric variables `sepal_length`,`sepal_width`,`petal_length`, and `petal_width	`, and 1 character variable `species`. There are `r sum(is.na(iris_with_missing))` missing values, 20 among each variable.

### Filling in missing values

A function `fill_na_function` is defined to fill `NA` according to the rules defined in the problem. The resulting dataframe after filling is shown below.

```{r}

fill_na_function = function(vec){
  if (is.numeric(vec)){
    mean = mean(vec,na.rm = TRUE)
    for (i in 1:length(vec)){
      if (is.na(vec[[i]])){
        vec = replace(vec,i,mean)
      }
    }
  }else if (is.character(vec)){
    for (i in 1:length(vec)){
      if (is.na(vec[[i]])){
        vec =  replace(vec,i,"virginica")
      }
    }
  }
  
  return(vec)
}

iris_without_missing = map_df(iris_with_missing,fill_na_function)

knitr::kable(iris_without_missing)
```






